\documentclass[10pt]{article}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{fullpage}
\usepackage{color}
\usepackage{bm}
\usepackage{url}
\usepackage{hyperref}
\usepackage{color,hyperref}
\definecolor{darkblue}{rgb}{0.0,0.0,0.3}
\hypersetup{colorlinks,breaklinks,
            linkcolor=darkblue,urlcolor=darkblue,
            anchorcolor=darkblue,citecolor=darkblue}

%Change to Helvetica
%\usepackage[scaled]{helvet}
%\renewcommand*\familydefault{\sfdefault} %% Only if the base font of the document is to be sans serif
%\usepackage[T1]{fontenc}

%Change to fourier
%\usepackage{fourier}
%\usepackage[T1]{fontenc} 

\usepackage[T1]{fontenc}
\usepackage{charter}
\usepackage[expert]{mathdesign}

\begin{document}
\noindent\textbf{Assessment Evaluation  - POST Math Prefresher - 9/2021 33 possible points } \\
We've just spent a magical week together. This is to see how much stuck from the course and to help you identify areas that could use more attention and development. Do your best.  \textbf{Note: THERE ARE TWO PAGES!} \\

\noindent \textbf{Background questions:} \\
\indent Name, discipline and subfield? \\

\vspace{3mm}
\noindent \textbf{Questions}

\begin{enumerate}

\item Basics
\begin{enumerate}
\item Explain the significance or use of the following symbols:
\begin{enumerate} 
\item $\Sigma$
\end{enumerate}


\item Solve: (no need to simplify but show steps/work if possible)
\begin{enumerate}
\item $5\leq x-3$ \\
\item $-9x-2>3$ \\
\item $|x-2|\leq3$ \\
\item $2e^{6x}=18$ \\
\item $e^{x^4}=1$ \\
\item $ln(x^2)=5$ \\
\item $\sum_{n=1}^{10} 2+5n $\\
\item $5!$ \\
\item $(\frac{x^6y^{-3}}{x^2y^3})^3$ \\
\end{enumerate}

\item Factor
\begin{enumerate}
\item $m^2+3m-4 $
\item $x^2+6x+9$
\item $2x^4-4x^2$
\end{enumerate}

\end{enumerate}

\item Set Theory
\begin{enumerate}
\item Explain the meaning of the following symbols:
\begin{enumerate}
\item $\in$
\item $\forall$
\end{enumerate}
\item Suppose $A=\{3, 6, 12\}$,  $B=\{\text{hat, bulldozer, forklift}\}$ and $C=\{x| x\text{ is a natural number}| x >3 \text{ and } x<11\}$
\begin{enumerate}
\item What is $A \cup B$?
\item Write the elements of C
\item What is $A \cap C$?
\item What is $A\setminus C$?
\end{enumerate}
\end{enumerate}



\item Functions \& Pre-Calculus
\begin{enumerate}
\item What is a continuous function?
\item Draw an increasing function. \\
\newline
\newline
\newline
\item What is a tangent line? What does it do? \\
\item What is a derivative? 
\end{enumerate}

\item Matrix Algebra
\begin{enumerate}
\item Give an example of a $3 \times 4$ matrix
\item Consider the following matrices: \\
\textbf{A} =
$
\begin{bmatrix}
  3 & 4 & 1 \\
  0 & 2 & 1 \\
\end{bmatrix}$
\textbf{B} =
$\begin{bmatrix}
  0 & 1 & 1 \\
  1 & 2 & 4 \\
\end{bmatrix}$
\textbf{C} =
$\begin{bmatrix}
  1 & 4 & 1 \\
  0 & 2 & 1 \\
  6 & 2 & 9 \\
\end{bmatrix}$ \\
\begin{enumerate}
\item Which matrices can be added together?  Why?
\item Add the matrices from the above response.  \\ \vspace{2mm}
\item Which matrices can be multiplied together?  Why?
\item Multiply the matrices from the above response. \\ \vspace{2mm}
\end{enumerate}


\end{enumerate}

\item Calculus 
\begin{enumerate}
\item what is the derivative of $4$? \\
\item what is the derivative of $2x$? \\
\item find the derivative of $7m^2-m+2$ \\
\item calculate the integral $\int_{0}^{5}  (x^3+0.5x^2+5x)dx$ \\
\item calculate the integral $\int e^x dx$ \\
\end{enumerate}



\end{enumerate} 



\end{document}