\documentclass[10pt]{article}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{fullpage}
\usepackage{color}
\usepackage{bm}
\usepackage{url}
\usepackage{hyperref}
\usepackage{color,hyperref}
\definecolor{darkblue}{rgb}{0.0,0.0,0.3}
\hypersetup{colorlinks,breaklinks,
            linkcolor=darkblue,urlcolor=darkblue,
            anchorcolor=darkblue,citecolor=darkblue}

%Change to Helvetica
%\usepackage[scaled]{helvet}
%\renewcommand*\familydefault{\sfdefault} %% Only if the base font of the document is to be sans serif
%\usepackage[T1]{fontenc}

%Change to fourier
%\usepackage{fourier}
%\usepackage[T1]{fontenc} 

\usepackage[T1]{fontenc}
\usepackage{charter}
\usepackage[expert]{mathdesign}

\begin{document}
\noindent\textbf{Pre-Assessment Evaluation  - Math Prefresher - 9/2021 34 possible points} \\
This is an assessment of your knowledge of topics we will cover in this course. You may know some or most of the answers, you may not. This will help me get to understand your background and how to focus our energies during our time together. Please do your best and don't stress if you're not sure.  \\\textbf{Note: THERE ARE TWO SIDES!} \\

\noindent \textbf{Background questions:} \\
\indent Name, discipline and subfield? \\
\vspace{3mm}

 What are you hoping to gain from this course? \\
\vspace{4mm}


\noindent \textbf{Questions}
\begin{enumerate}
\item Basics
\begin{enumerate}
\item Explain the significance or use of the following symbols:
\begin{enumerate} 
\item $\Pi$ (note: different from $\pi$)
\item $\Sigma$
\end{enumerate}


\item Solve: (no need to simplify but show steps/work if possible)
\begin{enumerate}
\item $4\geq x-7$ \\
\item $-9x+2>3$ \\
\item $|x-2|\leq2$ \\
\item $2e^{6x}=18$ \\
\item $e^{x^2}=1$ \\
\item $ln(x^2)=5$ \\
\item $\sum_{n=1}^{10} 3+n $\\
\item $4!$ \\
\item $(\frac{x^4y^{-3}}{x^2y^3})^3$ \\
\end{enumerate}

\item Factor
\begin{enumerate}
\item $m^2+3m+2 $
\item $x^2+5x+6$
\item $x^2+x$
\end{enumerate}

\end{enumerate}

\item Set Theory
\begin{enumerate}
\item Explain the meaning of the following symbols:
\begin{enumerate}
\item $\in$
\item $\forall$
\end{enumerate}
\item Suppose $A=\{3, 4, 5\}$,  $B=\{\text{hat, triangle, forklift}\}$ and $C=\{x| x\text{ is a natural number}| x >3 \text{ and } x<9\}$
\begin{enumerate}
\item What is $A \cup B$?
\item Write the elements of C
\item What is $A \cap C$?
\end{enumerate}
\end{enumerate}



\item Functions \& Pre-Calculus
\begin{enumerate}
\item What is a continuous function?
\item Draw an increasing function. \\
\newline
\newline
\newline
\item What is a tangent line? What does it do? \\
\end{enumerate}

\item Matrix Algebra
\begin{enumerate}

\item What experience do you have with matrix algebra?
\vspace{4mm}
\item Give an example of a $3 \times 4$ matrix: 
\vspace{4mm}
\end{enumerate}

\item Calculus 
\begin{enumerate}
\item what is the derivative of $4x$? \\
\item calculate the derivative of $3m^2-8m+5$ \\
\item calculate the integral $\int_{0}^{5}  (x^3+0.5x^2+5x)dx$ \\
\item calculate the integral $\int e^x dx$ \\
\end{enumerate}

\item Probability
\begin{enumerate}
\item What does $P(A)$ mean?
\item What is an independent event?
\vspace{3mm}
\item Define or explain Bayes' Rule:
\vspace{3mm}
\item What is a Venn diagram? Give, draw and label an example.
\end{enumerate}

\item Statistics
\begin{enumerate}
\item What is the difference between continuous and discrete variables?
\vspace{3mm}
\item What is a probability mass function? (define, explain, give an example)
\end{enumerate}

\end{enumerate} 



\end{document}