---
title: "Differentiation"
author: "[MACS 33000](https://jmclip.github.io/MACSS_math_camp/) <br /> University of Chicago"
output:
  xaringan::moon_reader
---

```{r child = here::here("_common-slides.Rmd"), cache = FALSE}

```

```{r packages, include=FALSE}
library(tidyverse)
library(patchwork)
library(here)
library(gganimate)
```

# Learning objectives

* Define the slope of a line
* Summarize tangent lines, rates of change, and derivatives
* Define derivative rules for common functions
* Apply the product, quotient, and chain rules for differentiation
* Summarize the exponential function and natural logarithms
* Identify properties of derivatives helpful to statistical methods

---

# What is calculus?

* Calculus
* Differential calculus
* Integral calculus
* Underlies most statistical and computational methods
* Crucial for **optimization**

--

## Potential uses

* Given data, what is the most likely value of a parameter(s)?
* Game theory: given another player's strategy, what is the action that maximizes utility?

---

# How functions change

* Derivatives
* A special limit
* Cover three broad ideas 
    1. Geometric interpretation/intuition
    1. Formulas/algebra derivatives
    1. Famous theorems

---

# The tangent as a limit

```{r tan-lines}
tan_fun <- function(x) x^2 - 3 * x + 3

tan_lines <- tibble(
  x0 = 2,
  y0 = tan_fun(x0),
  x1 = seq(from = 4, to = 2.001, by = -0.1),
  y1 = tan_fun(x1),
  slope = (y1 - y0) / (x1 - x0),
  intercept = y1 - slope * x1
) %>%
  mutate(id = row_number())

function_plot +
  stat_function(fun = tan_fun) +
  xlim(0, 5) +
  geom_abline(
    data = tan_lines,
    mapping = aes(
      intercept = intercept,
      slope = slope
    )
  ) +
  geom_point(
    data = tan_lines,
    mapping = aes(x = x0, y = y0)
  ) +
  geom_point(
    data = tan_lines,
    mapping = aes(x = x1, y = y1)
  ) +
  geom_text(
    data = tan_lines,
    mapping = aes(x = x0, y = y0, label = "P"), nudge_x = -0.1, nudge_y = 0.5, size = 7
  ) +
  geom_text(
    data = tan_lines,
    mapping = aes(x = x1, y = y1, label = "Q"), nudge_x = -0.1, nudge_y = 0.5, size = 7
  ) +
  # tangent line at x = 2
  geom_abline(slope = 1, intercept = -1, linetype = 2) +
  annotate(geom = "text", x = 4.5, y = 3, label = "T", size = 7) +
  labs(title = "Curve and tangent line") +
  transition_states(states = id, wrap = FALSE)
```

---

# The tangent as a limit

.pull-left[

```{r tan-lines-slim, ref.label = "tan-lines", fig.width = 6}

```

]

.pull-right[

$$
\begin{aligned}
P &= (x_0, y_0) \\
Q &= (x_1, y_1) \\
\text{slope of } L &= \frac{y_1 - y_0}{x_1 - x_0}
\end{aligned}
$$

]

---

# The tangent as a limit

.pull-left[

```{r tan-lines-slim2, ref.label = "tan-lines", fig.width = 6}
```

]

.pull-right[

$$
\begin{aligned}
P &= (x_0, y_0) \\
Q &= (x_1, y_1) \\
\text{slope of } L &= \frac{y_1 - y_0}{x_1 - x_0}
\end{aligned}
$$

### Substitute values

$$
\begin{aligned}
h &= x_1 - x_0 \\
x_1 &= x_0 + h \\
y_0 &= f(x_0) \\
y_1 &= f(x_0 + h) \\
\text{slope of } L &= \frac{f(x_0 + h) - f(x_0)}{h}
\end{aligned}
$$

]

---

# Derivative

Suppose $f:\Re \rightarrow \Re$. Measure rate of change at a point $x_{0}$ with a function $R(x)$,

$$R(x) = \frac{f(x) - f(x_{0}) }{ x- x_{0} }$$

* $R(x)$ defines the rate of change
* A derivative will examine what happens with a small perturbation at $x_{0}$

---

# Derivative

* Let $f:\Re \rightarrow \Re$. If the limit

    $$\begin{aligned} \lim_{x\rightarrow x_{0}} R(x) & = \frac{f(x) - f(x_{0}) }{x - x_{0}} \\ & = f^{'}(x_{0}) \end{aligned}$$

    exists then we say that $f$ is **differentiable** at $x_{0}$
* If $f^{'}(x_{0})$ exists for all $x \in \text{Domain}$, then we say that $f$ is differentiable

---

# Derivative

* Let $f$ be a function whose domain includes an open interval containing the point $x$
* The derivative of $f$ at $x$ is given by

    $$\frac{d}{dx}f(x) =\lim\limits_{h\to 0} \frac{f(x+h)-f(x)}{(x+h)-x} = \lim\limits_{h\to 0} \frac{f(x+h)-f(x)}{h}$$

--

## Derivative notation

* Leibniz notation: $\frac{d}{dx}(f(x))$
* Prime or Lagrange notation: $f'(x)$

---

# Example derivatives

```{r derivsimple, fig.cap = "The Derivative as a Slope", fig.height = 6}
fx <- function_plot +
  stat_function(fun = function(x) 2 * x, size = 0.5) +
  labs(
    title = expression(f(x) == 2 * x),
    y = expression(f(x))
  ) +
  expand_limits(y = 0)

fprimex <- function_plot +
  stat_function(fun = function(x) 2, size = 0.5) +
  labs(
    x = expression(x),
    y = expression(f ~ plain("'") ~ (x))
  )

gx <- function_plot +
  stat_function(fun = function(x) x^3, size = 0.5) +
  labs(y = expression(g(x)), title = expression(g(x) == x^3)) +
  expand_limits(y = 0)

gprimex <- function_plot +
  stat_function(fun = function(x) 3 * (x^2), size = 0.5) +
  labs(x = expression(x), y = expression(g ~ plain("'") ~ (x)))

fx + fprimex + gx + gprimex + plot_layout(ncol = 2)
```

* If $f'(x)$ exists at a point $x_0$, then $f$ is said to be **differentiable** at $x_0$
* Implies continuity of $f(x)$ at $x_0$

---

# Rates of change in a function

* Consider $y = f(x)$
* As $x$ changes from $x_0$ to $x_0 + h$, the value of the function changes from $f(x_0)$ to $f(x_0 + h)$
* Change in $x$ is $h$
* Change in $f(x)$ is $f(x_0 + h) - f(x_0)$
* Rate of change of $f(x)$ is defined to be

    $$\frac{f(x_0 + h) - f(x_0)}{h}$$

--

* Same thing as a derivative

---

# Rates of change in a function

```{r vote-spending, fig.height = 5}
vote_share <- function(x) 0.474 + 1e-8 * x - 5.05e-17 * x^2

function_plot +
  stat_function(fun = vote_share) +
  geom_abline(intercept = 0.60025, slope = 4.95e-9, color = "blue") +
  geom_vline(xintercept = 5e07, linetype = 2, color = "blue") +
  scale_x_continuous(labels = scales::dollar, limits = c(1e07, 1e08)) +
  scale_y_continuous(labels = scales::percent, limits = c(0.5, 1)) +
  labs(x = "Dollars spent in district",
       y = "Vote share")
```

--

* Rate of change $\leadsto$ return on vote share for dollars invested
* Instantaneous rate of change $\leadsto$ increase in vote share in response to infinitesimally small increase in spending
* A type of **limit**

---

# Examples of derivatives

* $f(x) = x^2$ and consider $x_{0} = 1$

$$
\begin{aligned}
\lim_{x\rightarrow 1}R(x) & = \lim_{x\rightarrow 1} \frac{x^2 - 1^2}{x - 1}  \\
 	& = \lim_{x\rightarrow 1} \frac{(x- 1)(x + 1) }{ x- 1}   \\
	& =  \lim_{x\rightarrow 1} x + 1  \\
	& = 2
\end{aligned}
$$

---

# Examples of derivatives

* $f(x) = |x|$ and consider $x_{0} = 0$

    $$\lim_{x\rightarrow 0} R(x) = \lim_{x\rightarrow 0} \frac{ |x| } {x}$$

--

* $\lim_{x \rightarrow 0^{-}} R(x) = -1$
* $\lim_{x \rightarrow 0^{+}} R(x) = 1$
* Not differentiable at $0$

---

# Continuity and derivatives

* $f(x) = |x|$ is **continuous** but not differentiable
* Change is **too abrupt**
* Suggests differentiability is a stronger condition

--

Let $f:\Re \rightarrow \Re$ be differentiable at $x_{0}$. Then $f$ is continuous at $x_{0}$.

---

# What goes wrong?

$$
\begin{aligned}
f(x)  & = x^{2} \text{ for all  } x \in \Re \setminus 0 \\
f(x) & = 1000  \text{ for  } x = 0
\end{aligned}
$$

* $f'(0)$

    $$\begin{aligned} \lim_{x \rightarrow 0 } R(x) & = \lim_{x \rightarrow 0 } \frac{f(x+h)-h}{ h   } \\ &= \lim_{x \rightarrow 0 }\frac{h^2}{h}\\ &= \lim_{x \rightarrow 0 } h \end{aligned}$$

* $\lim_{x \rightarrow 0 } h \neq 1000$, so the limit doesn't exist										

---

# Calculating derivatives

* Rarely will we take a limit to calculate derivative
* Rather, rely on **rules** and properties of derivatives
* **Important**: do not forget core intuition

--

## Strategy

* Algebra theorems
* Some specific derivatives
* Work on problems 

---

# Derivative rules: WRITE THESE DOWN

Suppose $a$ is some constant, $f(x)$ and $g(x)$ are functions:

$$
\begin{aligned}
f(x) &= x & \quad f^{'}(x) &= 1 \\
f(x) &= a x^{k} & \quad f^{'}(x) &= (a) (k) x ^{k-1} \\
f(x) &= e^{x } & \quad f^{'} (x) &= e^{x} \\
f(x) &= \sin(x) & \quad f^{'} (x) &= \cos (x) \\
f(x) &= \cos(x) & \quad f^{'} (x) &= - \sin(x) \\
\end{aligned}
$$

---

# Derivative rules

##### Constant rule $\left[k f(x)\right]' = k f'(x)$

##### Sum rule $\left[f(x)\pm g(x)\right]' = f'(x)\pm g'(x)$

##### Product rule $\left[f(x)g(x)\right]' = f'(x)g(x)+f(x)g'(x)$

##### Quotient rule $\frac{f(x)}{g(x)}' = \frac{f'(x)g(x)-f(x)g'(x)}{[g(x)]^2}, g(x)\neq 0$

##### Power rule $\left[x^k\right]' = k x^{k-1}$

---

# Derivatives with a friend

Differentiate the following functions:

1. $f(x)= x^3 + 5 x^2  + 4 x$, <span style="color: gray;"> at $x_{0} = 2$</span>
1. $f(x) = \sin(x) x^3$, <span style="color: gray;"> at $x_{0} = 2$</span>
1. $h(x) = \dfrac{e^{x}}{x^3}$, <span style="color: gray;"> at $x_{0} = 2$ </span>
1. $h(x) = \log (x) x^3$, <span style="color: gray;"> at $x_{0} = e$ </span>

---

# Composite functions

* Differentiating functions of functions

    $$x^2 + 1^2 \quad \text{vs.} \quad (x^2 + 1)^2$$

--

* Composite functions

    $$f \circ g=f[g(x)]$$

* Range of $g$ must be contained (at least in part) within the domain of $f$
* Domain of $f \circ g$ consists of all the points in the domain of $g$ for which $g(x)$ is in the domain of $f$

---

# Composite functions

$$
\begin{aligned}
f(x) &= \log x \in (0, \infty) \\
g(x) &= x^2 \in (-\infty, \infty)
\end{aligned}
$$

--

$$f[g(x)] = \log x^2 \in (-\infty, \infty) - \{0\}$$

$$g[f(x)] = [\log x]^2 \in (0, \infty)$$

---

# Chain rule

* $y = f[g(x)]$
* Derivative of $y$ with respect to $x$ is

    $$\frac{d}{dx} \{ f[g(x)] \} = f'[g(x)] g'(x)$$

* Derivative of the "outside" times the derivative of the "inside"
The chain rule can be thought of as the derivative of the "outside" times 
* "Outside" evaluated at the value of the inside function

---

# Chain rule

$$
\begin{aligned}
h(x) &= e^{2x} \\
g(x) &= e^{x} \\
f(x) &= 2x
\end{aligned}
$$

$$h(x) = g(f(x)) = g(2x) = e^{2x}$$

--

$$h^{'}(x) = g^{'}(f(x))f^{'}(x) = e^{2x}2$$

---

# Chain rule

$$
\begin{aligned}
h(x) &= \log(\cos(x) ) \\
g(x) &= \log(x) \\
f(x) &= \cos(x)
\end{aligned}
$$

$$h(x) = g(f(x)) = g( \cos(x)) = \log(\cos(x))$$

--

$$h^{'}(x) = g^{'}(f(x))f^{'}(x) = \frac{-1}{\cos(x)} \sin(x) = -\tan (x)$$

---

# Generalized power rule

* If $f(x)=[g(x)]^p$ for any rational number $p$

    $$f^\prime(x) =p[g(x)]^{p-1}g^\prime(x)$$

---

# Exponential function

```{r exp-func}
function_plot +
  stat_function(fun = exp) +
  xlim(-3, 3) +
  labs(title = expression(f(x) == e^x))
```


---

# Derivative of exponential function

$$\frac{d}{dx}(e^x) = e^x$$

* Why?

--

$$
\begin{aligned}
\frac{d}{dx}f(x) & = \lim\limits_{h\to 0} \frac{f(x+h)-f(x)}{h} \\
&= \lim\limits_{h\to 0} \frac{e^{x + h} - e^x}{h} \\
&= \lim\limits_{h\to 0} \frac{e^x e^h - e^x}{h} \\
&= \lim\limits_{h\to 0} \frac{e^x(e^h - 1)}{h} \\
&= e^x \lim\limits_{h\to 0} \frac{e^h - 1}{h}
\end{aligned}
$$

---

# Derivative of exponential function

```{r exp-limit}
function_plot +
  stat_function(fun = function(x) (exp(x) - 1) / x, size = 0.5) +
  xlim(-3, 3) +
  labs(
    title = expression(f(h) == frac(e^h - 1, h)),
    x = expression(h),
    y = expression(f(h))
  )
```

---

# Derivative of exponential function

$$
\begin{aligned}
\frac{d}{dx}f(x) & = \lim\limits_{h\to 0} \frac{f(x+h)-f(x)}{h} \\
&= \lim\limits_{h\to 0} \frac{e^{x + h} - e^x}{h} \\
&= \lim\limits_{h\to 0} \frac{e^x e^h - e^x}{h} \\
&= \lim\limits_{h\to 0} \frac{e^x(e^h - 1)}{h} \\
&= e^x \lim\limits_{h\to 0} \frac{e^h - 1}{h} \\
& = e^x (1) \\
& = e^x
\end{aligned}
$$

---

# Derivative of exponential function

```{r fig-derivexponent, fig.cap = "Derivative of the Exponential Function"}
{
  function_plot +
    stat_function(fun = function(x) exp(x), size = 0.5) +
    xlim(-3, 3) +
    labs(y = expression(f(x)))
} + {
  function_plot +
    stat_function(fun = function(x) exp(x), size = 0.5) +
    xlim(-3, 3) +
    labs(y = expression(f ~ plain("'") ~ (x)))
} +
  plot_annotation(expression(f(x) == e^x))
```

---

# Derivative of the natural logarithm

* Logarithm to base $e$ of $x$, where $e$ is defined as **Euler's number**

    $$y = \log_e (x) \iff x = e^y$$

* Relationship between $e^x$ and $\log_e(x)$

    $$\begin{aligned} e^{\log(x)} &= x \, \text{for every positive number} \, x \\ \log(e^y) &= y \, \text{for every real number} \, y \\ \end{aligned}$$

---

# Exponential function and natural logarithm

```{r exp-log, fig.cap = "Exponential function and natural logarithm"}
{
  function_plot +
    stat_function(fun = function(x) exp(x), size = 0.5, n = 1000) +
    xlim(-3, 3) +
    labs(
      title = "The exponential function",
      y = expression(e^x)
    )
} + {
  function_plot +
    stat_function(fun = function(x) log(x), size = 0.5, n = 1000) +
    xlim(0, 3) +
    labs(
      title = "Natural logarithm",
      y = expression(log(x))
    )
}
```

---

# Derivative of a natural logarithm
    
  $$(f^{-1})'(x) = \frac{1}{f'(f^{-1}(x))} $$
    
Let's find the derivative of $log(x)$. We already know the derivative of its inverse, $e^x$ is $e^x$. So, we begin with $f(x) = e^x$ (and $f'(x) = e^x$) and want to know $(f^{-1})'(x)$.
  
We know $f'(f^{-1}(x)) = f'(log(x)) = e^{log(x)} = x$
  
--
    

Therefore,
  $$\begin{aligned} (f^{-1})'(x) &= \frac{1}{f'(f^{-1}(x))} \\
   &= \frac{1}{(e^{log(x)})} \\
  &= \frac{1}{x} \end{aligned}$$

---

# Derivative of a natural logarithm

```{r fig-derivlog, fig.cap = "Derivative of the Natural Log"}
{
  function_plot +
    stat_function(fun = function(x) log(x), size = 0.5, n = 1000) +
    xlim(0, 3) +
    labs(y = expression(f(x)))
} + {
  function_plot +
    stat_function(fun = function(x) 1 / x, size = 0.5, n = 1000) +
    xlim(0, 3) +
    labs(y = expression(f ~ plain("'") ~ (x)))
} +
  plot_annotation(title = expression(f(x) == log(x)))
```

---

# Relevance of $e^x$ and $\log(x)$

* Growth over time (e.g. compounding interest)
* Elasticity models
* Variable transformation

---

# Derivatives and properties of functions

* Often used to **optimize** a function (tomorrow)
* But also reveal **average rates of change**
* Or crucial properties of functions

---

# Rolle's theorem

> Suppose $f:[a, b] \rightarrow \Re$.  Suppose $f$ has a relative maxima or minima on $(a,b)$ and call that $c \in (a, b)$.  Then $f'(c) = 0$.

--

```{r rolles-theorem, fig.height = 7}
function_plot +
  stat_function(aes(color = "f(x)"), fun = function(x) -1 * (x^2) + 4) +
  geom_hline(aes(color = "f'(x)", yintercept = 4), linetype = 2) +
  xlim(-3, 3) +
  scale_color_brewer(type = "qual", guide = guide_legend(reverse = TRUE)) +
  labs(
    title = "Rolle's theorem",
    y = expression(f(x)),
    color = NULL
  ) +
  theme(legend.position = "bottom")
```

---

# Rolle's theorem

> Suppose $f:[a, b] \rightarrow \Re$.  Suppose $f$ has a relative maxima or minima on $(a,b)$ and call that $c \in (a, b)$.  Then $f'(c) = 0$.

* Consider (without loss of generalization) a relative maximum $c$
* Consider the left-hand and right-hand limits

    $$\begin{aligned} \lim_{x \rightarrow c^{-}} \frac{f(x) - f(c) }{x - c } & \geq 0  \\ \lim_{x \rightarrow c^{+}} \frac{f(x) - f(c) } {x - c }  & \leq 0 \end{aligned}$$

* But we also know that 

    $$\begin{aligned} \lim_{x \rightarrow c^{-}} \frac{f(x) - f(c ) }{x - c } & = f^{'}(c)  \\ \lim_{x \rightarrow c^{+}} \frac{f(x) - f(c) } {x - c }  &  =  f^{'}(c) \end{aligned}$$


---

# Rolle's theorem

* The only way, then, that we can have something $\geq 0$ equal to $\leq 0$ is if $f^{'}(c) = 0$ :
$$\lim_{x \rightarrow c^{-}} \frac{f(x) - f(c) }{x -c}  = \lim_{x \rightarrow c^{+}} \frac{f(x) - f(c) } {x - c}$$ 

---

# Mean value theorem

* If $f:[a,b] \rightarrow \Re$ is continuous on $[a,b]$ and differentiable on $(a,b)$, then there is a $c \in (a,b)$ such that 

    $$f^{'}(c) = \frac{f(b) - f(a) } { b - a}$$

--

* Rolle's theorem is a special case of the MVT, where $f'(c) = 0$

---

# Mean value theorem

```{r mean-value-theorem}
# function for MVT demo
mvt_fun <- function(x) -x^2 + 5

# calculate chord points
a <- -4
b <- 2
f_a <- mvt_fun(a)
f_b <- mvt_fun(b)
chord_points <- tribble(
  ~x, ~y, ~label,
  a, f_a, "a, f(a)",
  b, f_b, "b, f(b)"
)

# calculate tangent line parallel to chord
f_c_slope <- (f_b - f_a) / (b - a)
f_c_int <- 6

# draw plot
ggplot(
  data = tibble(x = 0),
  mapping = aes(x = x)
) +
  stat_function(fun = mvt_fun) +
  xlim(-4, 2) +
  ylim(NA, 10) +
  geom_abline(intercept = f_c_int, slope = f_c_slope, color = "blue") +
  geom_segment(
    mapping = aes(
      x = a,
      y = f_a,
      xend = b,
      yend = f_b
    ),
    linetype = 2
  ) +
  geom_point(data = chord_points, mapping = aes(x, y)) +
  geom_text(data = chord_points, mapping = aes(x, y, label = label), nudge_x = 0, nudge_y = -1, size = 7) +
  labs(
    title = "Mean value theorem",
    x = expression(x),
    y = expression(f(x))
  ) +
  annotate(
    geom = "text", x = -3, y = 3, label = expression(f ~ plain("'") ~ (c) == frac(f(b) - f(a), b - a)),
    color = "blue", size = 7
  )
```

---

# Applications of the mean value theorem

* This will come up in a formal theory article. You'll at least know where to look
* It allows us to say lots of powerful stuff about functions
* Especially useful for approximating derivatives (see next set of slides)

---

# Applications of the mean value theorem

Suppose that $f:[a,b] \rightarrow \Re$ is continuous on $[a,b]$ and differentiable on $(a,b).$ Then, 

1. If $f^{'}(x) \neq 0$ for all $x \in (a,b)$ then $f$ is 1-1
1. If $f^{'}(x) = 0$ then $f(x)$ is constant 
1. If $f^{'}(x)> 0$ for all $x \in (a,b)$ then then $f$ is strictly increasing
1. If $f^{'}(x)<0$ for all $x \in (a,b)$ then $f$ is strictly decreasing

---

# If $f^{'}(x) \neq 0$ for all $x \in (a,b)$ then $f$ is 1-1

```{r one-to-one}
{
  function_plot +
    stat_function(fun = function(x) x) +
    ggtitle("One-to-one function")
} + {
  function_plot +
    stat_function(fun = function(x) x^2) +
    ggtitle("Not a one-to-one function")
}
```

---

# If $f^{'}(x) \neq 0$ for all $x \in (a,b)$ then $f$ is 1-1

* Suppose that $f$ is not 1-1. Then there is $x, y \in (a,b)$ such that $f(x) = f(y)$. Then, 

    $$f'(c) = \frac{f(x) - f(y)}{x- y} = \frac{0}{x -y}  = 0$$

* This means $f' \neq 0$ for all $x$
* **Contradiction**

---

# If $f^{'}(x) = 0$ then $f(x)$ is constant

* By way of contradiction, suppose that there is $x, y \in (a,b)$ such that $f(x) \neq f(y)$. But then, 

    $$f'(c) = \frac{f(x) - f(y) } {x - y} \neq 0$$

---

# If $f^{'}(x)> 0$ for all $x \in (a,b)$ then $f$ is strictly $\uparrow$

* By way of contradiction, suppose that there is $x, y \in (a,b)$ with $y<x$ but $f(y)>f(x)$. But then, 

    $$f'(c) = \frac{f(x) - f(y) }{x - y } < 0,$$ which contradicts $f^{'}(x)> 0$
* Proof for strictly decreasing is the reverse of this

---

# Cauchy mean value theorem

* Suppose $f$ and $g$ are differentiable functions and $a$ and $b$ are real numbers such that $a < b$
* Suppose also that $g'(x) \neq 0$ for all $x$ such that $a < x < b$
* There exists a real number $c$ such that $a < c < b$ and

    $$\frac{f'(c)}{g'(c)} = \frac{f(b) - f(a)}{g(b) - g(a)}$$

--

* The classical mean value theorem is the special case where $g(x) = x$ for all $x$

---

# Implications and a little help from Cauchy

Suppose we want to find

$$\lim_{x \rightarrow a} \frac{f(x)}{g(x)}$$

where $f$ and $g$ are continuous functions

* Suppose $g(a) \neq 0$, then we can do this:

    $$\lim_{x \rightarrow a} \frac{f(x)}{g(x)} = \frac{f(a)}{g(a)}$$

* BUT if $f(a) = g(a) = 0$... OH NO! What do we do with an indeterminate form (e.g. $\frac{0}{0}$)?

---

# New bestie: L'Hôpital's Rule

* Suppose that $f(a) = g(a) = 0$ and $g'(x) \neq 0$ if $x$ is close but not equal to $a$ (also applies for $\frac{\pm\infty}{\infty}$).
* Then

    $$\lim_{x \rightarrow a} \frac{f(x)}{g(x)} = \lim_{x \rightarrow a} \frac{f'(x)}{g'(x)}$$

--

$$
\begin{aligned}
\frac{f(x) - f(a)}{g(x) - g(a)} &= \frac{f'(c)}{g'(c)} \\
\frac{f(x) - 0}{g(x) - 0} &= \frac{f'(c)}{g'(c)} \\
\frac{f(x)}{g(x)} &= \frac{f'(c)}{g'(c)}
\end{aligned}
$$

---

# Example of L'Hôpital's Rule

.pull-left[

```{r lhopital-1, fig.width = 6}
function_plot +
  stat_function(fun = function(x) ((1 + x)^(1/3) - 1) / (x - x^2)) +
  xlim(-.5, .5) +
  labs(title = expression(f(x) == frac((1 + x)^(1/3) - 1, x - x^2)),
       y = expression(f(x)))
```

]

.pull-right[

$$
\begin{aligned}
f'(x) &= \frac{1}{3} (1 + x)^{-2/3} \\
f'(0) &= \frac{1}{3} (1)^{-2/3} = \frac{1}{3} (1) = \frac{1}{3} \\
g'(x) &= 1 - 2x \\
g'(0) &= 1 - 2(0) = 1
\end{aligned}
$$

$$\lim_{x \rightarrow a} \frac{f(x)}{g(x)} = \lim_{x \rightarrow a} \frac{f'(x)}{g'(x)} = \frac{1/3}{1} = \frac{1}{3}$$

]

---

# Example of L'Hôpital's Rule

.pull-left[

```{r lhopital-2, fig.width = 6}
function_plot +
  stat_function(fun = function(x) (x - log(1 + x)) / x^2, n = 201) +
  xlim(-1, 1) +
  labs(title = expression(f(x) == frac(x - log(1 + x), x^2)),
       y = expression(f(x)))
```

]

.pull-right[

$$
\begin{aligned}
f(x) &= x - \log(1 + x) \\
f'(x) &= 1 - \frac{1}{1 + x} \\
g(x) &= x^2 \\
g'(x) &= 2x
\end{aligned}
$$

]

---

# Simplify the expression

$$
\begin{aligned}
L &= \lim_{x \rightarrow 0} \frac{1 - \frac{1}{1 + x}}{2x} \\
&= \lim_{x \rightarrow 0} \frac{1}{2x} - \frac{\frac{1}{1 + x}}{2x} = \lim_{x \rightarrow 0} \frac{1}{2x} - \frac{1}{2x(1 + x)} \\
&= \lim_{x \rightarrow 0} \frac{1(1 + x)}{2x(1 + x)} - \frac{1}{2x(1 + x)} \\
&= \lim_{x \rightarrow 0} \frac{1(1 + x) - 1}{2x(1 + x)} \\
&= \lim_{x \rightarrow 0} \frac{1 + x - 1}{2x(1 + x)} \\
&= \lim_{x \rightarrow 0} \frac{x}{2x(1 + x)} = \lim_{x \rightarrow 0} \frac{1}{2(1 + x)} \\
&= \lim_{x \rightarrow 0} \frac{1}{2(1 + 0)} = \frac{1}{2}
\end{aligned}
$$

---

# Iterative application

$$\lim_{x \rightarrow a} \frac{f(x)}{g(x)} = \lim_{x \rightarrow a} \frac{f'(x)}{g'(x)} = \lim_{x \rightarrow a} \frac{f''(x)}{g''(x)} = \ldots$$

$$
\begin{aligned}
f''(x) &= \frac{1}{(1 +x)^{2}} \\
g''(x) &= 2 \\
\lim_{x \rightarrow 0} \frac{f''(x)}{g''(x)} &= \frac{(1 + x)^{-2}}{2} = \frac{1^{-2}}{2} = \frac{1}{2}
\end{aligned}
$$

---

class: inverse, center, middle

# RECAP

---
RECAP:

* Good chunk of calc 1
* Rate of change
* Intuition for derivative
* Mathematical foundation for derivative
* Rules for derivatives
* Making magic: MVT + implications

---

class: inverse, center, middle

## Helpful formulas

---

# Derivative rules: WRITE THESE DOWN

Suppose $a$ is some constant, $f(x)$ and $g(x)$ are functions:

$$
\begin{aligned}
f(x) &= x & \quad f^{'}(x) &= 1 \\
f(x) &= a x^{k} & \quad f^{'}(x) &= (a) (k) x ^{k-1} \\
f(x) &= e^{x } & \quad f^{'} (x) &= e^{x} \\
f(x) &= \sin(x) & \quad f^{'} (x) &= \cos (x) \\
f(x) &= \cos(x) & \quad f^{'} (x) &= - \sin(x) \\
\end{aligned}
$$

---

# Derivative rules

##### Constant rule $\left[k f(x)\right]' = k f'(x)$

##### Sum rule $\left[f(x)\pm g(x)\right]' = f'(x)\pm g'(x)$

##### Product rule $\left[f(x)g(x)\right]' = f'(x)g(x)+f(x)g'(x)$

##### Quotient rule $\frac{f(x)}{g(x)}' = \frac{f'(x)g(x)-f(x)g'(x)}{[g(x)]^2}, g(x)\neq 0$

##### Power rule $\left[x^k\right]' = k x^{k-1}$
