---
title: "Integration and integral calculus"
author: "[MACS 33000](https://jmclip.github.io/MACSS_math_camp/) <br /> University of Chicago"
output:
  xaringan::moon_reader
---

```{r child = here::here("_common-slides.Rmd"), cache = FALSE}

```

```{r packages, include=FALSE}
library(tidyverse)
library(patchwork)
library(here)
library(gganimate)
library(plotly)
```

# Learning objectives

* Summarize areas, slices, and integrals
* Apply common rules to calculate definite integrals
* Define the fundamental theorem of calculus
* Calculate antiderivatives and integrals using integration by substitution and integration by parts
* Evaluate improper integrals
* Evaluate multivariate integrals

---

# Indefinite integration

* Calculate the derivative $f=F'$ of a function $F$
* Reverse operation: find the function $F$ for which $f$ is its derivative
* Antiderivative

---

# Indefinite integration

$$f(x) = \frac{1}{x^2}$$

$$
\begin{aligned}
\int \frac{1}{x^2} dx &= -\frac{1}{x} + c \\
\frac{d}{dx} \left[ -\frac{1}{x} + c \right] &= \frac{1}{x^2}
\end{aligned}
$$

$$F(x) = \int f(x) \,dx$$

---

# Many possible antiderivatives

$$
\begin{aligned}
f(x) &= x^2-4 \\
F(x) &= \int\limits f(x) \,dx
\end{aligned}
$$

--

```{r integralc, fig.height = 6}
fx <- function_plot +
  stat_function(fun = function(x) x^2 - 4, size = 0.5) +
  labs(y = expression(f(x)))

Fx <- function_plot +
  stat_function(fun = function(x) (x^3) / 3 - 4 * x, size = 0.5, linetype = "dashed") +
  stat_function(fun = function(x) (x^3) / 3 - 4 * x + 1, size = 0.5, linetype = "dotted") +
  stat_function(fun = function(x) (x^3) / 3 - 4 * x - 1, size = 0.5, linetype = "dotdash") +
  labs(y = expression(integral(f(x) * dx)))

fx + Fx + plot_layout(ncol = 1)
```

* Multiple antiderivatives for an arbitrary constant $c$

---

# Common rules of integration

1. Constants are allowed to slip out: $\int a f(x)\,dx = a\int f(x)\,dx$
1. Integration of the sum is sum of integrations:  $\int [f(x)+g(x)]\,dx=\int f(x)\,dx + \int g(x)\,dx$
1. Reverse Power-rule:  $\int x^n \,dx = \frac{1}{n+1} x^{n+1} + c$
1. Exponents are still exponents:  $\int e^x \,dx = e^x +c$
1. Recall the derivative of $\log(x)$ is one over $x$, and so:   $\int \frac{1}{x} \,dx = \log x + c$
1. Reverse chain-rule:  $\int e^{f(x)}f^\prime(x)\,dx = e^{f(x)}+c$
1. More generally:  $\int [f(x)]^n f'(x)\,dx = \frac{1}{n+1}[f(x)]^{n+1}+c$
1. Remember the derivative of a log of a function:  $\int \frac{f^\prime(x)}{f(x)}\,dx=\log f(x) + c$

---

# Simplify the indefinite integrals
  
1. $\int 3x^2 \,dx$
1. $\int (2x+1) \,dx$
1. $\int e^x e^{e^x} \,dx$

---

# Definite integral

* Indefinite integral $\leadsto$ definite integral
* Different objective: find the area under a function

--

* Determine the area $A(R)$ of a region $R$ defined by a curve $f(x)$ and some interval $a\le x \le b$
* Approximation with rectangles
    * Divide the interval $a\le x\le b$ into $n$ subintervals of length $\Delta x$
    * Base of each rectangle is $\Delta x$
    * Height is $f(x)$ at the midpoint of that interval

    $$A(R) \approx S(f,\Delta x)=\sum\limits_{i=1}^n f(x_i)\Delta x$$

* Riemann sum

---

# Riemann sum

$$A(R)=\lim\limits_{\Delta x\to 0}\sum\limits_{i=1}^n f(x_i)\Delta x$$

```{r defintfig, fig.cap = "The Riemann Integral as a Sum of Evaluations"}
# generate intervals
f3 <- function(x) -15 * (x - 5) + (x - 5)^3 + 50

d <- tibble(seq_int = seq(from = 0.1, to = 1, by = .1)) %>%
  mutate(seq = map(seq_int, ~ seq(from = 0, to = 10, by = .x)),
         d = map(seq, ~ tibble(x = .x,
                               f = f3(.x)))) %>%
  unnest(d)

ggplot(data = d, mapping = aes(x, f, width = seq_int)) +
  geom_bar(stat = "identity", position = "identity",
           fill = "gray", alpha = 0.5, color = "black") +
  stat_function(fun = f3) +
  scale_x_continuous(expand = c(0, 0)) +
  labs(title = "Riemann integral as a sum of evaluations",
       subtitle = expression(f(x) == -15 * (x - 5) + (x - 5)^3 + 50),
       x = expression(x),
       y = expression(f(x))) +
  transition_time(time = seq_int, range = c(1, .1)) +
  ease_aes("cubic-in-out")
```

---

# Definite integral

* Riemann integral of $f$ from $a$ to $b$

    $$\lim\limits_{\Delta x\to 0} \sum\limits_{i=1}^n f(x_i)\Delta x = \int\limits_a^b f(x) \,dx$$

* Suppose $f:[a,b] \rightarrow \Re$ is a continuous function. Then $f$ is integrable
* Suppose $f:[a,b]\rightarrow \Re$ is a monotonic function.  Then $f$ is integrable

---

# Counterexamples

* Suppose $f:[0,1]\rightarrow \frac{1}{x}$

    $$\int_{0}^{1} \frac{1}{x} \,dx$$

* $\frac{1}{x}$ is not integrable on $[a,b]$

--

----------------------------

$$\begin{aligned}f(x) &= 1 \text{ if } x \text{ rational}  \\	 &= 0 \text{ if } x \text{ irrational} \end{aligned}$$

* Not integrable - discontinuity

---

# Fundamental theorem of calculus

* Riemann sum = antiderivative
* Function $f$ bounded on $[a,b]$ and continuous on $(a,b)$
* $F(x)$ denotes the definite integral from $a$ to $x$:

    $$F(x)=\int\limits_a^x f(t) \,dt, \quad a\le x\le b$$

* $F(x)$ has a derivative at each point in $(a,b)$
    
    $$F^\prime(x)=f(x), \quad a<x<b$$
    
* Differentiation is the inverse of integration

---

# Fundamental theorem of calculus

* Calculating indefinite integrals
* Function $f$ bounded on $[a,b]$ and continuous on $(a,b)$
* $F$ is any function that is continuous on $[a,b]$ such that $F'(x)=f(x)$ on $(a,b)$. Then

    $$\int\limits_a^bf(x)\,dx = F(b)-F(a)$$

--

### Estimation procedure for definite integral

$$\int\limits_a^b f(x)\,dx$$

1. Find the indefinite integral $F(x)$
2. Evaluate $F(b)-F(a)$

---

# Uniform distribution

Suppose $f:\Re \rightarrow \Re$, with 

$$
\begin{aligned}
f(x) &=  1 \text{ if  } x \in [0,1]  \\
f(x) &= 0 \text{ otherwise }
\end{aligned}
$$

--

What is the area under $f(x)$ from $[0, 1/2]$?

--

$$\begin{aligned}\int_{0}^{1/2}  f(x)\,dx &= \int_{0}^{1/2} 1 \,dx \\	&= x|_{0}^{1/2}  \\	&= (1/2) - (0 )  \\							&= 1/2 \end{aligned}$$

---

# Area under a line

Suppose $f:\Re \rightarrow \Re$, with 

$$f(x) = x$$

--

Evaluate the $\int_{2}^{t}f(x)\,dx$

--

$$\begin{aligned}\int_{2}^{t}f(x)\,dx &= \int_{2}^{t} x \,dx  \\	&= \frac{x^{2} }{2} |_{2}^{t}   \\		&= \frac{t^2}{2} - \frac{2^2}{2}  \\	&= \frac{t^2}{2} - \frac{4}{2} = \frac{t^2}{2} - 2\end{aligned}$$

---

# Area under a curve

Solve $\int\limits_1^3 3x^2 \,dx$

$$\begin{aligned}f(x) &= 3x^2 \\F(x) &= x^3 + c \\\int\limits_1^3 3x^2 \,dx &= F(3) - F(1) \\&= (3^3 + c) - (1^3 + c) \\&= 27 + c - 1 - c \\&= 26\end{aligned}$$

---

# Common rules for definite integrals

1. There is no area below a point

    $$\int_a^a f(x)\,dx=0$$
    
1. Reversing the limits changes the sign of the integral

    $$\int_a^b f(x)\,dx=-\int_b^a f(x)\,dx$$
    
1. Sums can be separated into their own integrals

    $$\int_a^b [\alpha f(x)+\beta g(x)]\,dx = \alpha \int_a^b f(x)\,dx + \beta \int_a^b g(x)\,dx$$
    
1. Areas can be combined as long as limits are linked

    $$\int_a^b f(x) \,dx +\int_b^c f(x)\,dx = \int_a^c f(x)\,dx$$

---

# Challenge problems

1. $\int\limits_1^1 3x^2 \,dx$
1. $\int\limits_0^4 (2x+1) \,dx$
1. $\int\limits_{-2}^0 e^x e^{e^x} dx + \int\limits_0^2 e^x e^{e^x} \,dx$

---

# Integration by substitution

* Tricky integrands
* Alternative methods
* Related to the chain rule

--

$$\int g(x)\,dx$$

* Generate new function $u(x)$ such that

    $$g(x)=f[u(x)]u'(x)$$

* Antiderivative of $f$ as $F$

    $$\frac{d}{dx} F[u(x)]=f[u(x)]u'(x)$$

    $$\int g(x) \,dx= \int f[u(x)]u'(x)\,dx = \int \frac{d}{dx} F[u(x)]\,dx = F[u(x)]+c$$

---

# Integration by substitution

1. Identify some part of $g(x)$ that might be simplified by substituting in a single variable $u$ (which will then be a function of $x$)
1. Determine if $g(x)\,dx$ can be reformulated in terms of $u$ and $du$
1. Solve the indefinite integral
1. Substitute back in for $x$

---

# Example

$$\int \frac{x}{x^2 + 1} \,dx$$

--

Reexpress as

$$\frac{1}{2} \int \frac{2x}{x^2 + 1} \,dx$$

--

$$\begin{aligned}u &= x^2 + 1 \\du &= 2x \,dx \\\int \frac{x}{x^2 + 1} \,dx &= \frac{1}{2} \int \frac{2x}{x^2 + 1} \,dx \\&= \frac{1}{2} \int \frac{1}{u} \,du \\&= \frac{1}{2} \log(u) + c \\&= \frac{1}{2} \log(x^2 + 1) + c \\\end{aligned}$$

---

# Integration by parts

* Related to the product rule of differentiation

    $$\begin{aligned}\frac{d}{dx} f(x) g(x) &=  f(x) g'(x) + g(x) f'(x) \\\frac{d}{dx}(uv) &= u\frac{dv}{dx}+v\frac{du}{dx}\end{aligned}$$

* Integrating this and rearranging

    $$\int u\frac{dv}{dx}\,dx= u v - \int v \frac{du}{dx}\,dx$$

    $$\int u(x) v'(x)\,dx=u(x)v(x) - \int v(x)u'(x)\,dx$$

    $$\int u \,dv = u v - \int v \,du$$

    where $du=u'(x)\,dx$ and $dv=v'(x)\,dx$

* Find expressions for $u$ and $dv$ that yield an expression that's more easily evaluated

---

# Integration by parts

$$\int x \log(x) \,dx$$

--

$$
\begin{aligned}
u &= \log(x) & \quad dv & = x \, dx \\
du &= \frac{1}{x} \,dx & \quad v &= \frac{1}{2}x^2
\end{aligned}
$$

--

$$
\begin{aligned}
\int x \log(x) \, dx &= u v - \int v \,du \\
				&= \log(x) \times \frac{1}{2} x^2 - \int (\frac{1}{2} x^2) (\frac{1}{x} \, dx) \\
				&= \frac{1}{2} x^2 \log(x) - \int \frac{1}{2}x \,dx \\
				&= \frac{1}{2} x^2 \log(x) - \frac{1}{2} \int x \,dx \\
				&= \frac{1}{2} x^2 \log(x) - \frac{1}{2} \left( \frac{1}{2}x^2 \right) + c \\
				&= \frac{1}{2} x^2 \log(x) - \frac{1}{4} x^2 + c \\
\end{aligned}
$$

---

# Infinite integrals

$$
\begin{aligned}
\int_1^X x^{-2} \,dx &= -\frac{1}{x} |_1^X \\
&= -\frac{1}{X} - -\frac{1}{1} \\
&= -\frac{1}{X} + 1 \\
&= 1 - X^{-1}
\end{aligned}
$$

--

As $X \rightarrow \infty, \frac{1}{X} \rightarrow 0$

$$\lim_{X \rightarrow \infty} \int_1^X x^{-2} \,dx = 1$$

$$\int_1^\infty x^{-2} \,dx = 1$$

---

# Infinite integrals

```{r int-x-2, fig.caption = "The area is equal to 1"}
function_plot +
  stat_function(fun = function(x) 1 / x^2, n = 501) +
  geom_area(stat = "function", fun = function(x) 1 / x^2, fill = "blue",
            xlim = c(1, 5)) +
  xlim(0, 5) +
  ylim(0, 10) +
  labs(title = expression(y == x^-2))
```

---

# Infinite integrals

* Over an infinite domain
* Still has a finite value for the definite integral

--

## Convergence of integrals

* Continuous function $f(x)$ defined for all $x \geq a$
* Integral $\int_a^X f(x)\,dx$ approaches a finite limit $L$ as $X \rightarrow \infty$

    $$\int_a^\infty f(x)\,dx = L$$

    * Indefinite integral **converges**
* If $\int_a^X f(x)\,dx$ does not approach a finite limit as $X \rightarrow \infty$, then infinite integral **diverges**

---

# Diverging integral

$$f(x) = \frac{1}{x}$$

--

$$\begin{aligned}\int_{1}^{\infty} \frac{1}{x} \,dx &= \lim_{t \rightarrow \infty} \int_{1}^{t} \frac{1}{x}  \,dx \\										&= \lim_{t \rightarrow \infty} (\log x)|_{1}^{t} \\										&= \lim_{t \rightarrow \infty} (\log t) - \lim_{t \rightarrow \infty} (\log 1) \end{aligned}$$

---

# Two-sided infinite integrals

If $f(x)$ is a continuous function defined for all $x \leq a$, and if the definite integral $\int_Y^a f(x) \,dx$ approaches a finite limit as $Y \rightarrow -\infty$, we denote the limit by

$$\int_{-\infty}^a f(x) \,dx$$

--

If the integrals $\int_{-\infty}^a f(x) \,dx$ and $\int_a^\infty f(x) \,dx$ both exist, we denote their sum by

$$\int_{-\infty}^\infty f(x) \,dx$$

--

* Extremely important for probability distributions

    $$\int_{-\infty}^\infty f(x) \,dx = 1$$

---

# Improper integrals

* Integral where the integrand is not defined at one of the limits of integration

    $$I = \int_0^1 \frac{1}{\sqrt{x}} \,dx$$

    * $\frac{1}{\sqrt{x}} \rightarrow \infty$ as $x \rightarrow 0$

--

* Define and evaluate $I$ as

    $$I = \lim_{\delta \downarrow0} \int_\delta^1 \frac{1}{\sqrt{x}} \,dx$$

    $$\begin{aligned}I &= \lim_{\delta \downarrow0} \int_\delta^1 \frac{1}{\sqrt{x}} \,dx \\&= \lim_{\delta \downarrow0} \left[ 2\sqrt{x} \right]_\delta^1 \\&= \lim_{\delta \downarrow0} (2 - 2\sqrt{\delta}) \\&= 2\end{aligned}$$

---

# Improper integrals

* Function $f(x)$ is defined and continuous for $a < x \leq b$, but is not defined for $x = a$
* If the integral

    $$\int_{a + \delta}^b f(x) \,dx$$

    tends to a finite limit $I$ as $\delta \downarrow 0$, integral $\int_a^b f(x) \,dx$ exists and has the value $I$
* Otherwise, integral $\int_a^b f(x) \,dx$ diverges

---

# Monte Carlo and integration

* Compute some integral $\int_{-\infty}^{\infty} xf(x) \,dx$
* $f(x)$ is complicated

    $$f(x) = \frac{\exp\left(- \frac{(x- \mu)^2}{2\sigma^2} \right) }{\sqrt{2\pi}}$$

--

## Monte Carlo simulation

* Generate random draws from $f(x)$:

    $$d_{1}, d_{2}, d_{3}, ..., d_{T}$$

* Approximation

    $$\text{Expected Value} = \frac{1}{T}\sum_{i =1}^{T} d_{i}$$

    * As $T \rightarrow \infty$, $\text{Expected value} \rightarrow \int_{-\infty}^{\infty} xf(x) \,dx$

---

# Monte Carlo simulation

```{r monte-carlo}
rerun(.n = 10, tibble(x = rnorm(1000))) %>%
  bind_rows(.id = "id") %>%
  group_by(id) %>%
  mutate(x_bar = cummean(x),
         n_id = row_number()) %>%
  ggplot(mapping = aes(x = n_id, y = x_bar, color = factor(id))) +
  geom_line() +
  scale_color_brewer(type = "qual", palette = "Set3") +
  labs(x = "Number of draws",
       y = "Estimate") +
  theme(legend.position = "none") +
  transition_reveal(along = n_id)
```

---

# Multivariate integration

* Function $f:X \rightarrow \Re^{1}$, with $X \subset \Re^{2}$
* Integrate a function over an area

---

# Multivariate integration

$$A = \{x, y : x \in[0,1], y \in [0,1] \}$$

```{r a-2d-rect}
function_plot +
  geom_rect(data = tribble(
  ~"x1", ~"y1", ~"x2", ~"y2",
  0, 0, 1, 1
), mapping = aes(x = NULL,
                 xmin = x1, xmax = x2,
                 ymin = y1, ymax = y2)) +
  xlim(-1, 1) +
  ylim(-1, 1)
```

---

# Multivariate integration

$$A = \{x, y: x^2 + y^2 \leq 1 \}$$

```{r a-2d-circle}
function_plot +
  ggforce::geom_circle(aes(x = NULL, x0 = 0, y0 = 0, r = 1), fill = "grey") +
  xlim(-1, 1) +
  ylim(-1, 1) +
  coord_fixed()
```

---

# Multivariate integration

$$A = \{ x, y: x< y, x, y \in (0,2) \}$$

```{r a-2d-triangle}
function_plot +
  geom_polygon(data = tribble(
  ~"x", ~"y",
  0, 0,
  2, 0,
  2, 2
), mapping = aes(x = x, y = y)) +
  xlim(-1, 2) +
  ylim(-1, 2)
```

---

# Multivariate integration

* $f:X \rightarrow \Re$ where $X \subset \Re^{n}$
* $f$ is integrable over $A \subset X$ if we are able to calculate its area with refined partitions of $A$
* Write the integral $I =\int_{A} f(\boldsymbol{x}) d\boldsymbol{A}$

---

# Fubini's theorem

* Suppose $A = [a_{1}, b_{1}]  \times [a_{2}, b_{2} ] \times \ldots \times [a_{n}, b_{n}]$
* $f:A \rightarrow \Re$ is integrable

    $$\int_{A} f(\boldsymbol{x}) d\boldsymbol{A} = \int_{a_{n}}^{b_{n}} \int_{a_{n-1}}^{b_{n-1}} \ldots \int_{a_{2}}^{b_{2}} \int_{a_{1}}^{b_{1}} f(\boldsymbol{x})\,dx_{1} \,dx_{2} \ldots \,dx_{n-1} \,dx_{n}$$

* Start with the inside integral $x_{1}$ is the variable, everything else a constant
* Work inside to out, iterating
* At the last step, we should arrive at a number

---

# Multivariate uniform distribution

* Suppose $f:[0,1] \times [0,1] \rightarrow \Re$ and $f(x_{1}, x_{2}) = 1$ for all $x_{1}, x_{2} \in [0,1]\times[0,1]$
* What is $\int_{0}^{1}\int_{0}^{1} f(x) \,dx_{1} \,dx_{2}$?

--

$$\begin{aligned}\int_{0}^{1}\int_{0}^{1} f(x) \,dx_{1} \,dx_{2} &= \int_{0}^{1} \int_{0}^{1} 1 \,dx_{1} \,dx_{2} \nonumber \\															&= \int_{0}^{1} x_{1}|_{0}^{1} \,dx_{2} \\															&= \int_{0}^{1} (1 - 0) \,dx_{2}  \\															&= \int_{0}^{1} 1 \,dx_{2}  \\															&= x_{2}|_{0}^{1}  \\															&= 1 \end{aligned}$$

---

# Another example

* Suppose $f:[a_{1}, b_{1} ] \times [a_{2}, b_{2} ]  \rightarrow \Re$ is given by

    $$f(x_{1}, x_{2} ) = x_{1} x_{2}$$

* Find $\int_{a_{2}}^{b_{2}} \int_{a_{1}}^{b_{1}} f(x_{1}, x_{2} )\,dx_{1} \,dx_{2}$

--

$$\begin{aligned}\int_{a_{2}}^{b_{2}} \int_{a_{1}}^{b_{1}} f(x_{1}, x_{2} )\,dx_{1} \,dx_{2} &= \int_{a_{2}}^{b_{2}} \int_{a_{1}}^{b_{1}} x_{1} x_{2} \,dx_{1} \,dx_{2} \\ &= \int_{a_{2}}^{b_{2}} \frac{x_{1}^2}{2} x_{2} |_{a_{1}}^{b_{1}} \,dx_{2} \\ &= \frac{b_{1}^{2} - a_{1}^{2} }{2}  \int_{a_{2}}^{b_{2}} x_{2} \,dx_{2} \\ &= \frac{b_{1}^{2} - a_{1}^{2} }{2} \left( \frac{x_{2}^{2} }{2} |_{a_{2}}^{b_{2}} \right ) \\  &= \frac{b_{1}^{2} - a_{1}^{2} }{2} \frac{b_{2}^{2} - a_{2}^{2} }{2}\end{aligned}$$

---

# More complicated bounds of integration

```{r a-2d-triangle}
```

---

# Example: more complicated region

* Suppose $f:[0,1] \times [0,1] \rightarrow \Re$, $f(x_{1}, x_{2}) = x_{1} + x_{2}$. Find area of function where $x_{1} < x_{2}$.

--

* Determine the bound - if $x_{1}< x_{2}$, $x_{1}$ can take on any value from $0$ to $x_{2}$

--

$$\begin{aligned}\iint_{x_{1}< x_{2}}  f(\boldsymbol{x}) &= \int_{0}^{1} \int_{0}^{x_{2}} x_{1} + x_{2} \,dx_{1} \,dx_{2}  \\&= \int_{0}^{1} x_{2} x_{1} |_{0}^{x_{2}} \,dx_{2}  + \int_{0}^{1} \frac{x_{1}^{2} }{2} |_{0}^{x_{2} }  \\&= \int_{0}^{1} x_{2}^{2} \,dx_{2}  + \int_{0}^{1} \frac{x_{2}^2}{2}  \\&= \frac{x_{2}^{3} }{3}|_{0}^{1} + \frac{x_{2}^{3}}{6}|_{0}^{1}  \\&= \frac{1}{3}  + \frac{1}{6}  \\&= \frac{3}{6} = \frac{1}{2}  \end{aligned}$$

---

# Switch the bounds

$$\begin{aligned}\iint_{x_{1}<x_{2}} f(\boldsymbol{x}) &= \int_{0}^{1} \int_{x_{1}}^{1} x_{1} + x_{2} \,dx_{2} \,dx_{1} \\													&= \int_{0}^{1} x_{1}x_{2}|_{x_{1}}^{1}  +  \int_{0}^{1} \frac{x_{2}^{2}}{2} |_{x_{1}}^{1}\,dx_{1}  \\													&= \int_{0}^{1} x_{1} - x_{1}^2 + \int_{0}^{1} \frac{1}{2} - \frac{x_{1}^2}{2} \,dx_{1}  \\													&=  \frac{x_{1}^2}{2}|_{0}^{1}  - \frac{x_{1}^{3}}{3}|_{0}^{1}  + \frac{x_{1}}{2}|_{0}^{1}  - \frac{x_{1}^{3}}{6}|_{0}^{1}  \\													&= \frac{1}{2}  - \frac{1}{3}  + \frac{1}{2} - \frac{1}{6}  \\													&= 1 - \frac{3}{6}  \\													&= \frac{1}{2}  \end{aligned}$$
